---
title: "INFSCI 2595 Final Project"
subtitle: "Part 1A, EDA: Distribution"
author: "Zhenyu Li"
date: '2022-03-29'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load_packages, message=FALSE, warning=FALSE}
library(tidyverse)
source("./scripts/utils.R")
theme_set(theme_linedraw())
```

----

In section 1A I will explore the data for basic properties and distribution of the variables.  

- Check data summary, missing values, range, unique values, etc for each variable. 
- Examine distribution of individual variables. 
- Check correlation between continuous inputs. 

**Read in the data:**

```{r message=FALSE, warning=FALSE}
df <- readr::read_csv("./data/final_project_train.csv", col_names = TRUE)
```

**Examine the dataset by `glimpse`.**

```{r}
df %>% glimpse()
```

**Select numeric and non-numeric fields into respective variables for easy reference later on.**  

```{r}
sel_num <- df %>% select_if(is_double) %>% select(!rowid) %>% select(sort(names(.))) %>% colnames()
sel_cat <- df %>% select(!one_of(sel_num)) %>% select(!rowid) %>% colnames()
```

**Perform required data transformations for the `response` and `outcome` fields.**  

```{r}
df_eda <- df %>% mutate(
  log_response = log(response),
  binary_outcome = ifelse(outcome=="event", 1, 0))
```

----

## Categorical Variables

### Data Summary

**Get a summary of the categorical variables `region` and `customer`, as well as the binary response `outcome`. Check missing values, unique values. **

```{r}
summarize_cate_var <- function(df) {
  tibble(
    variable = names(df),
    n_missing = map_dbl(df, ~sum(is.na(.))),
    n_levels = map_dbl(df, n_distinct),
  ) %>% mutate(
    percent_missing = n_missing/nrow(df)*100,
  )
}

summ_cate <- df_eda %>% select(all_of(sel_cat)) %>% summarize_cate_var()
summ_cate %>% knitr::kable(caption = "categorical variables overview")
```

- There is no missing value in categorical variables.
- There are 3 levels for `region` and 9 levels for `customer`.


### Data Distribution

**Count the unique levels for each category and sort by count for each category.**

```{r}
df_eda %>% select(sel_cat) %>% pivot_longer(everything(), names_to ="var", values_to="level") %>% 
  count(var,level) %>% arrange(var,n) %>% 
  knitr::kable(caption = "categorical variables count")
```

**Use barchart to visualize the distribution of categorical input and binary response variables.**

```{r out.width=600, fig.width=6, fig.asp=0.5, warning=FALSE}
df_eda %>% select(sel_cat) %>% 
  pivot_longer(everything()) %>%
  ggplot() +
  geom_bar(aes(x=value, fill=value)) +
  facet_grid(~name, scales="free_x", space="free_x") +
  xlab("") + theme(legend.position="none")
```

- The categorical inputs
  - For `region` variable, there are more sale attempts for customers in region `ZZ`, than `YY`, than `XX`.
  - For `customer` variable, besides `Other` group which is a pooled category, there are more sale attempts to customer `G`, followed by customer `M`.

- The binary outcome
  - The `outcome` needs to be {0,1} encoded for logistic regression. [preprocessing]
  - The `outcome` variable is unbalanced: non-event (meet sale goal) consist of larger portion of the overall observations. The observed overall event probability is about 0.188. [interpretation]

```{r}
df_eda %>% select(binary_outcome) %>% colMeans()
```

----

## Continuous Variables

### Data Summary

**Get a high level summary of the numeric variables, check missing values, unique values, range and median.**

```{r}
summarize_nume_var <- function(df) {
  tibble(
    variable = names(df),
    n_missing = map_dbl(df, ~sum(is.na(.))),
    n_unique = map_dbl(df, n_distinct),
    min = map_dbl(df, min),
    median = map_dbl(df, median),
    max = map_dbl(df, max)
  ) %>% mutate(
    percent_missing = n_missing/nrow(df)*100,
    percent_unique = n_unique/nrow(df)*100,
  )
}

summ_nume <- summarize_nume_var(df %>% select(all_of(sel_num)))
summ_nume %>% knitr::kable(caption = "numeric variables overview")
```

- There is no missing value in all numeric variables.
- All sentiment derived features are numeric.
  - They are at different scales, and need to be standardized. [preprocessing]
  - Most are clearly continuous, but some may be discrete (very limited unique values).

**Examine the variables with less than 5% unique values.**

```{r}
cols_tmp <- summ_nume %>% filter(percent_unique < 5) %>% select("variable") %>%  as_vector() %>% unname()

df_eda %>% select(all_of(cols_tmp)) %>% lapply(unique) %>% lapply(sort)
```

- They look like categorical variables with numeric levels. Because of the large number of categories, I'll still treat them as continuous.  

### Data Distribution

**Examine the data distribution of the input variables. **

```{r out.width=1200, fig.width=12, fig.asp=0.65, warning=FALSE}
p <- df_eda %>% add_column(xs_98=NA, xs_99=NA) %>%
  pivot_longer(starts_with("x")) %>%
  ggplot(aes(x=value)) +
    geom_density(aes(color=region), bins=50) +
    # geom_histogram(bins=50) +
    facet_wrap(~name, scales="free", ncol=8) +
    ylab("") + xlab("")
plot_grid(p)
```

**Check if inputs are normally distributed using q-q plot**

```{r out.width=1200, fig.width=12, fig.asp=0.65, warning=FALSE}
p <- df_eda %>% add_column(xs_98=NA, xs_99=NA) %>% select(starts_with("x")) %>%
  pivot_longer(everything()) %>%
  ggplot() +
    geom_qq(aes(sample=value), size=0.2, alpha=0.1) +
    facet_wrap(~name, scales="free", ncol=8) +
    xlab("")+ylab("")
plot_grid(p)
```

- Most features look normally distributed, except:
  - The `xw` that corresponds to word count data, which is lower-bounded by 0. Two of them `xw_02`, `xw_03` look quite skewed.
  - The `xs_05` looks like it also has 0 as lower bound.
- Inputs are not uniform across the `region` category. 

**Examine the distribution of the continuous response.**

```{r out.width=500, fig.width=5, fig.asp=0.65, warning=FALSE}
df_eda %>% select(response) %>%
  ggplot() +
    geom_histogram(aes(x=response), bins=50)
```

- The continuous `response` is lower-bounded by 0, and is not normally distributed with the long tail. Must be log-transformed. [preprocessing]
- Majority of the `response` are below 5, with a tail extending beyond 10. There is also one data point with over 20 hours of interaction time. This may be an outlier that worth consider removing. [preprocessing???]

**Visualize q-q plot of log-transformed response.**

```{r out.width=300, fig.width=5, fig.asp=1, warning=FALSE}
df_eda %>%
  ggplot() +
  geom_qq(aes(sample=log_response)) 
```


**Check the data points with over 10 hours of sale time. **

```{r}
df_eda %>% filter(response > 10) %>% arrange(desc(response)) %>% knitr::kable(caption = "potential outliers")
```


### Conditioned on Categorical

**Examine continuous input summary statistics based on region. **

```{r out.width=1200, fig.width=12, fig.asp=0.60, warning=FALSE}
p <- df_eda %>% add_column(xs_98=NA, xs_99=NA) %>%
  pivot_longer(starts_with("x")) %>%
  ggplot(aes(y=value, x=region, color=region)) +
  geom_boxplot(outlier.size = 0.2) +
  facet_wrap(~name, scales="free", ncol=8) +
  xlab("") + ylab("")
plot_grid(p)
```


**Examine continuous input summary statistics based on customer. **

```{r out.width=1200, fig.width=12, fig.asp=0.60, warning=FALSE}
p <- df_eda %>% add_column(xs_98=NA, xs_99=NA) %>%
  pivot_longer(starts_with("x")) %>%
  ggplot(aes(y=value, x=customer, color=customer)) +
  geom_boxplot(outlier.size = 0.2) +
  facet_wrap(~name, scales="free", ncol=8) +
  xlab("") + ylab("")
plot_grid(p)
```


----

## Input Correlation

Input correlation could lead to uncertainty in posterior model coefficients and difficulty when using Lasso regularized regression.  

### Continuous Input

**Check correlation between continuous input variables.** 

```{r out.width=500, fig.width=5, fig.asp=1, warning=FALSE}
df_eda %>% select(all_of(sel_num)) %>% select(-response) %>%
  cor() %>%
  corrplot::corrplot(type="upper", method="color", diag=F)
```

- There are clear correlation patterns between some input variables.
  - The sequential blue diagonal squares indicate the positive correlations between features with same feature index among the lexicon derived features `xa`,`xb` and `xn`.
  - The many 4-cell blue-red diagonals squares indicate correlation between feature index `02`,`03` and between index `05`,`06` features for each set of inputs. 
  - These are more obvious when reordered. 

**Reorder the correlation plot to view the correlated terms. **

```{r out.width=500, fig.width=5, fig.asp=1, warning=FALSE}
df_eda %>% select(all_of(sel_num)) %>% select(-response) %>%
  cor() %>%
  corrplot::corrplot(
    type="upper", 
    method="color", 
    diag = F,
    order="hclust",
    hclust.method="ward.D2")
```

